### 2024-03-10 chapter_3/src/chapter_3.qmd
working on the models a little bit

---
title: |
  How does visual information influence social interactions?
code-repo: |
  For replication, go to: <https://github.com/DamonCharlesRoberts/book_project/chapter_3>.
author:
  - name: Damon C. Roberts
    email: damon.roberts-1@colorado.edu
    orcid: 0000-0002-4360-3675
    title: PhD Candidate
    affiliations:
      - id: CU
        name: University of Colorado Boulder
        department: Political Science
        address: 333 UCB
        city: Boulder
        region: CO 
        postal-code: 80309-0333
abstract: |
  If colors have associations with partisan groups, do they shape our willingness to converse with others? This chapter argues that colors are not only relevant to politics and act as a reliable heuristic for people to determine the partisanship of another person, but it additionally makes the argument that this has downstream consequences for our attitudes and behaviors. The existing literature on political deliberation is currently grappling with the capacity of conversations to persuade others about politics. There is mixed evidence suggesting that it can get people to change their minds under the right conditions.  However, these "right" conditions are often challenging to develop organically. This chapter presents a theory that I refer to as the snap-judgment model of political visual information processing that explains one potential source of this limitation. Here, I present a pre-registered report that elaborates on this literature, how the theory helps with this puzzle, and presents two research designs test my claims. I conclude the pre-registered report with a simulation exploring the sensitivity of conclusions I can draw from such a research design and resulting modeling strategies.

bibliography: ["../../../assets/deliberation-chapter-references.bib"]
format:
  hikmah-pdf:
    # Spit out in drafts directory
    latex-output-dir: "../../out/prr"
    # Use biblatex-chicago
    biblatex-chicago: true
    biblio-style: authordate
    biblatexoptions: 
      - backend=biber
      - autolang=hyphen
      - isbn=false
      - uniquename=false
    geometry:
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
      - heightrounded
    linestretch: 2
    fontsize: 12pt
date: today
nocite: |
  @allaire_quarto_2022
execute:
  echo: false
  warning: false
  message: false
  cache: false
params:
  replicate: false
---
{{< pagebreak >}}

# Introduction

@@

# The cognitive architecture of attitudes

@@

# The cognitive architecture and color in politics

@@

# Study 1 {#sec-study-1}

## Study design

@@

## Modeling strategy

@@

# Study 2

## Study design

@@

# Simulation {#sec-simulation}

The stated models above have theoretical motivations. Unfortunately for the social scientist, our theoretical motivations do not necessarily mean that our inferential ones also follow. The purpose of these simulations is not to convince myself or the reader that my stated modeling strategy is appropriate and "good" in some objective sense. Rather, its purpose is to give me a sense of the sensitivity of my ability to draw correct conclusions despite the reasonable amounts of uncertainty that quantitative work generates.

I intend to fit these models through the Bayesian regression framework. There are a number of benefits to Bayesian approaches to regression and hypothesis testing that scholars have debated for centuries. I am not going to hash out those debates here. However, from the perspective of pre-registering my expectations of a model, a Bayesian approach to regression seems particularly appropriate. The primary reason for this is that the Bayesian approach to regression requires that you report your prior beliefs about the explanatory effect of your predictor variables on an outcome. You incorporate these prior beliefs into your model as a distribution that indicates what you think is the most likely size of the effect and the probability of other less-likely effect sizes. This allows you to reflect uncertainty about the size of the effect but still makes explicit what you believe. You multiply your prior distribution with your likelihood distribution constructed with the data. This approach enables me to report my beliefs and my associated uncertainty about those effect sizes. They do not dictate the model, however. The more data you have, the less weight your prior beliefs have -- even ones that indicate you have little uncertainty about your beliefs. As the size of the data removes the "subjectivity" that many believe to be a problem introduced in Bayesian approaches to regression, one particularly important part of this pre-registered report is to demonstrate the sensitivity of my ability to draw inferences based on the size of my sample Additionally, some of my stated models are more complex than others. Therefore, the amount of information my model needs to be sure that my hypotheses are correct increases.

I first start by using the `fabricatr` package [@blair_fabricatr_2022] to simulate population data for both studies. I generate these simulated data from data generating processes that I define.^[I include the particulars of these data generating processes in the supplementary materials. I do not include them here because the primary goal of this exercise is to see whether or not I can draw correct inferences about reasonably similar data generating processes rather than on me defining these data generating processes as correct themselves.] I randomly draw from this population dataset at four different sample sizes: 200, 400, 600, and 800. For each of these sample sizes, I draw 100 samples. I do not perform these simulations for [Study 2](#study-2) as the design is largely the same, as well as the modeling strategies.

```{r}
#| label: setup
#setwd("./src/prr")
# modularly load packages
box::use(
  ./R/sample[
    study_simulator
  ]
  , ./R/model[
    fit_sims
  ]
  , ./R/plot[...]
  , data.table[...]
  , parallel[...]
  , brms[
    make_stancode
    , brmsformula
    , set_prior
    , bernoulli
    , cumulative
  ]
  , rstan[
    stan_model
  ]
)
# Set up cores
cores <- detectCores() - 1
# List sample size
list_sample_size <- c(200, 400, 600, 800)
# Number of samples
int_n_samples <- 1000
```

```{r}
#| label: simulate
if (params$replicate == TRUE) {
  # sample the data at different sample sizes
  list_simulated <- mclapply(
    list_sample_size
    , function(x) {
      df_sample <- study_simulator(
        study = 1
        , n = x
        , n_samples = int_n_samples
        , seed = 121022
      )
      # Return df_sample
      return(df_sample)
    }
    , mc.cores = cores
  )
  # store in a rdata file
  save(
    list_simulated
    , file = "../../data/temp/prr/simulated_samples.RData"
  )
} else {
  load(
    file = "../../data/temp/prr/simulated_samples.RData"
  )
}
```

On each of these samples, I fit the stated models using the `rstan` package [@stan_development_team_rstan_2023]. For all models, I define my prior distribution as a normal distribution with a mean of 0 and a standard deviation of 1. While this is not truly my prior beliefs of the effect sizes, I am using the prior distribution as an informed prior that states that I largely believe the effects to be zero, or at least to be very small. This acts as a quite conservative test of my models' ability to lead me to the correct inferences. Once fitted, I find the minimum and maximum value on the posterior distribution for each of my coefficients and identify whether that posterior distribution overlaps with zero. My hypotheses are directional in that I not only believe that the effect of my predictor terms should not be zero, but I also believe that these effects are either positive or negative.

I calculate a true positive rate by first ensuring that the minimum and maximum values of my posterior distribution for each $\hat{\beta_i}$ coefficient do not overlap with zero. If I would come to the wrong conclusion about my hypothesis by examining my posterior distribution, I record a value of 0; if it is correct, I code it as a 1. I then average over the 100 samples for each sample size. This gives me a true positive rate for each sample size for each $\hat{\beta_i}$ coefficient.

```{r}
#| label: make-stan-code
if (params$replicate == TRUE) {
  df_example <- list_simulated[[1]][
    sample == 1,
  ]
  df_example_split_rep <- df_example[
    party_id == "Strong Republican" | party_id == "Republican" | party_id == "Lean Republican"
  ]
  df_example_split_dem <- df_example[
    party_id == "Strong Democrat" | party_id == "Democrat" | party_id == "Lean Democrat"
  ]
  # Model 1: Predict whether they notice
  formula_study_1_model_1 <- brmsformula(
    notice ~ prime
  )
  prior_study_1_model_1 <- set_prior(
    "normal(0,1)"
    , class = "b"
  )
  stancode_study_1_model_1 <- make_stancode(
    formula = formula_study_1_model_1
    , data = df_example
    , family = bernoulli(link = "logit")
    , prior = prior_study_1_model_1
    , save_model = "./STAN/study_1_model_1.stan"
  )

  # Model 2: Predict willingness
    #* Full interaction
  formula_study_1_model_2 <- brmsformula(
    willing ~ blue_shirt + prime + party_id + blue_shirt * prime * party_id + gender + education + income + conflict + attention
  )
  prior_study_1_model_2 <- set_prior(
    "normal(0,1)"
    , class = "b"
  )
  stancode_study_1_model_2 <- make_stancode(
    formula = formula_study_1_model_2
    , data = df_example
    , family = cumulative(link = "logit")
    , prior = prior_study_1_model_2
    , save_model = "./STAN/study_1_model_2.stan"
  )
    #* Split sample
  formula_study_1_model_2_alt <- brmsformula(
    willing ~ blue_shirt + prime + blue_shirt * prime + gender + education + income + conflict + attention
  )
  prior_study_1_model_2_alt <- set_prior(
    "normal(0, 1)"
    , class = "b"
  )
  stancode_study_1_model_2_alt <- make_stancode(
    formula = formula_study_1_model_2_alt
    , data = df_example_split_rep
    , data2 = df_example_split_dem
    , family = cumulative(link = "logit")
    , prior = prior_study_1_model_2_alt
    , save_model = "./STAN/study_1_model_2_alt.stan"
  )
}
```

```{r}
#| label: fit-simulated-models
if (params$replicate == TRUE) {
  # Model 1
    #* Compile the stan model
  compiled_study_1_model_1 <- stan_model(
    file = "./STAN/study_1_model_1.stan"
    , model_name = "study_1_model_1"
  )
    #* Define model formula
  formula_study_1_model_1 <- brmsformula(
    notice ~ prime
  )
    #* Define the family
  family_study_1_model_1 <- bernoulli(link = "logit")
    #* Define priors
  prior_study_1_model_1 <- set_prior(
    "normal(0,1)"
    , class = "b"
  )
    #* Fit model at different sample sizes
  list_fitted_samples <- mclapply(
    list_simulated
    , function(x) {
      fit_sims(
        data_frame = x
        , compiled_stan_model = compiled_study_1_model_1
        , formula = formula_study_1_model_1
        , family = family_study_1_model_1
        , prior = prior_study_1_model_1
      )
    }
    , mc.cores = cores
  )
    #* collapse into a single data.table object
  names(list_fitted_samples) <- list_sample_size
  df_fitted_samples_model_1 <- rbindlist(
    list_fitted_samples
    , idcol = "sample_size"
  )[
    , sample_size := factor(sample_size)
  ]
  # Model 2
    #* Compile the stan model
  compiled_study_1_model_2 <- stan_model(
    file = "./STAN/study_1_model_2.stan"
    , model_name = "study_1_model_2"
  )
    #* Define model formula
  formula_study_1_model_2 <- brmsformula(
    willing ~ blue_shirt + prime + party_id + blue_shirt * prime * party_id + gender + education + income + conflict + attention
  )
    #* Define the family
  family_study_1_model_2 <- cumulative(link = "logit")
    #* Define priors
  prior_study_1_model_2 <- set_prior(
    "normal(0, 1)"
    , class = "b"
  )
    #* Fit model at different sample sizes
  list_fitted_samples_model_2 <- mclapply(
    list_simulated
    , function(x) {
      fit_sims(
        data_frame = x
        , compiled_stan_model = compiled_study_1_model_2,
        , parameters = c("b[1]", "b[2]", "b[3]", "b[4]")
        , formula = formula_study_1_model_2
        , priors = prior_study_1_model_2
        , family = family_study_1_model_2
      )
    }
    , mc.cores = cores
  )
    #* collapse into a single data.table object
  names(list_fitted_samples_model_2) <- list_sample_size
  df_fitted_samples_model_2 <- rbindlist(
    list_fitted_samples_model_2
    , idcol = "sample_size"
  )[
    , sample_size := factor(sample_size)
  ]
    #* alternative model
      #** compile model
  compiled_study_1_model_2_alt <- stan_model(
    file = "./STAN/study_1_model_2_alt.stan"
    , model_name = "study_1_model_2_alt"
  )
      #** formula
  formula_study_1_model_2_alt <- brmsformula(
    willing ~ blue_shirt + prime + blue_shirt * prime
  )
      #** clean data
  list_sim_rep_data <- lapply(
    list_simulated
    , function(x) {
      x[
        party_id == "Strong Republican" | party_id == "Republican" | party_id == "Lean Republican"
      ]
    }
  )
  list_sim_dem_data <- lapply(
    list_simulated
    , function(x) {
      x[
        party_id == "Strong Democrat" | party_id == "Democrat" | party_id == "Lean Democrat"
      ]
    }
  )
      #** fit model
  list_fitted_samples_model_2_alt <- mclapply(
    list_simulated
    , function(x) {
        fit_sims(
          data_frame = x
          , compiled_stan_model = compiled_study_1_model_2_alt
          , parameters = c("b[1]", "b[2]", "b[3]")
          , formula = formula_study_1_model_2_alt
          , priors = prior_study_1_model_2_alt
          , family = family_study_1_model_2
        )
    }
    , mc.cores = cores
  )
  names(list_fitted_samples_model_2_alt) <- list_sample_size
  df_fitted_samples_model_2_alt <- rbindlist(
    list_fitted_samples_model_2_alt
    , idcol= "sample_size"
  )

  # Save results
    #* study 1, model 1
  save(
    df_fitted_samples_model_1
    , file = "../../data/temp/prr/fitted_model_1.RData"
  )
    #* study 1, model 2
  save(
    df_fitted_samples_model_2
    , file = "../../data/temp/prr/fitted_model_2.RData"
  )
    #* study 1, model 2 alternate
  save(
    df_fitted_samples_model_2_alt
    , file = "../../data/temp/prr/fitted_model_2_alt.RData"
  )
} else {
  load(
    file = "../../data/temp/prr/fitted_model_1.RData"
  )
  load(
    file = "../../data/temp/prr/fitted_model_2.RData"
  )
  load(
    file = "../../data/temp/prr/fitted_model_2_alt.RData"
  )
}
```

The following histograms report the true positive rate for each $\hat{\beta_i}$ coefficient at different sample sizes for [Study 1](#study-1). Across models and across the two studies, I see that most models perform pretty well at any of the sample sizes. It is important to note that these sample sizes reflect the number of observations included in the models, not the total sample size. As many online convenience samples struggle with inattentive respondents, insincere responses, bots, etcetera [see @kennedy_strategies_2021], I am going to be quite conservative and try to recruit at least 800 respondents for these two studies.

```{r}
#| label: fig-true-positive-study-1-model-1
#| fig-cap: True positive rate from simulations - Model 1
#| fig-width: 8
#| layout-nrow: 1
plot_study_1_model_1 <- true_positive_plot(data_frame = df_fitted_samples_model_1)
plot_study_1_model_1[[1]]

```{r}
#| label: fig-true-positive-study-1-model-2
#| fig-cap: True positive rate from simulations - Model 2
#| fig-width: 8
#| layout-nrow: 2
#| fig-subcap:
#|   - $\hat{\beta}_1$
#|   - $\hat{\beta}_2$
#|   - $\hat{\beta}_3$
#|   - $\hat{\beta}_4$
plot_study_1_model_2 <- true_positive_plot(data_frame = df_fitted_samples_model_2)
plot_study_1_model_2[[1]]
plot_study_1_model_2[[2]]
plot_study_1_model_2[[3]]
plot_study_1_model_2[[4]]
```

```{r}
#| label: fig-true-positive-study-1-model-2-alt
#| fig-cap: True positive rate from simulations - Model 2, Split sample
#| fig-width: 8
#| layout-nrow: 2
#| fig-subcap:
#|   - $\hat{\beta}_1$
#|   - $\hat{\beta}_2$
#|   - $\hat{\beta}_3$

plot_study_1_model_2_alt <- true_positive_plot(data_frame = df_fitted_samples_model_2_alt)
plot_study_1_model_2_alt[[1]]
plot_study_1_model_2_alt[[2]]
plot_study_1_model_2_alt[[3]]
```

# Discussion

Previous chapters provide (as of right now) preliminary evidence supporting my claim that color matters and convey group-based partisan information to average citizens. This chapter pushes the envelope in a few ways. First, this chapter seeks to move past the argument that "people notice color, and it is relevant in politics." Specifically, this chapter seeks to examine whether it has the capacity to shape attitudes and behavior by elaborating on the mechanisms by which it does so. Second, it examines the limitations or the scope conditions of this claim. It argues that this works so long as politics is salient. Third, this chapter argues that color does not simply act as an alternative form of political information influencing political attitudes and their resulting behaviors but that it shapes the way we consider more traditionally considered forms of political information, such as stated policy positions. Fourth, this chapter explores the way in which that color influences our tendencies to avoid conversations about politics with out-partisans and to be more amenable to conversations with those we already agree with.

This pre-registered report documents my expectations before collecting or seeing any data that I can leverage to explore these claims. It also provides documentation of my concrete hypotheses, modeling strategies, and expectations about the sensitivity by which my research design and modeling strategies allow me to draw correct inferences from the data.

>>>>

# How does visual information influence social interactions? {#sec-3}

```{r}
#| echo: false

# Load the functions
box::use(
    datasummary[...]
)

## Introduction

@@

## The cognitive architecture of attitudes

@@

## The cognitive architecture and color in politics

@@

## Study 1 {#sec-study-1}

### Study design

@@

### Modeling strategy

@@

## Study 2

### Study design

@@
